---
title: "Assignment2_draft"
author: "Shiyuan Zhang"
date: "2023-06-07"
output: 
  bookdown::html_document2:
  fig_caption: yes
---

```{r}
install.packages("rnaturalearth")
install.packages("dslabs")
install.packages("rnaturalearthdata")

```


```{r LoadLibrary}
library(tidyverse)
library(plotly)
library(bookdown)
library(rnaturalearth)
library(rnaturalearthdata)
library(maps)
library(dslabs)
library(knitr)
library(scales)
library(dplyr)
library(gridExtra)
library(ISLR)
library(MASS)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(ggmap)
library(mapdata) 
library(kableExtra)
library(countrycode)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r LoadData2}
#Load merged data
data_init <- read.csv("Alcohol/Merge_all_2010_2020_updated.csv")

#Subset dataset for the section
data_init_mod <- data_init %>%
  select(c("Entity", "Year", "Code", "Death_alcohol_use_disorders"))

```


*Jane's Section* 


```{r, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}

# read in data set 
data = read.csv('Merge_all_2010_2020_updated.csv')

```

```{r, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}

# extracting columns to use from data 
extractedData =  data[, c(1, 2, 3, 12)]

# omit NA values
newExtractedData = na.omit(extractedData)

```

**Introduction**
Alcohol usage is a common occurrence in today's society, however the consumption amount varies across geographical regions and can be influenced by different factors. The report explores the relationship between total alcohol consumption and the geographical regions, as well as the significant influence of religion in alcohol usage globally.   


**Conclusion** 
Total alcohol consumption presented diverging trends over time, as some regions experienced growth while others declined.  Particularly, the Middle East, North Africa, and the Greater Indonesia region displayed consistency in low alcohol usage, which suggests a potential influence of religion given the prevalence of Islamic values in these countries. However, further information is needed to determine the relationship between religion and alcohol consumption. 


**Reference** 
H, Ritchie and M, Roser. (2018) <i>Alcohol Consumption</i>. OurWorldInData.org. https://ourworldindata.org/alcohol-consumption' 


*Question 1*

-   How has the total alcohol consumption per capita changed over time in relation to geographical regions?

**Analysis**

Table \@ref(tab:Table1) presented the overall median global growth of alcohol consumption from 2000 to 2018. However, Figure \@ref(fig:Figure1) revealed that regions were still experiencing diverging patterns during this period. Regions such as 'East Asia and Pacific', 'South Asia', 'North America' and 'Middle East and North Africa' increased in alcohol consumption, while other regions like 'Europe and Central Asia', 'Latin America and Caribbean' and 'Sub-Saharan Africa' experienced declines.

Table \@ref(tab:Table1) provided insights of the average (6.18L) and median (6.11L) alcohol consumption in 2010, this year stands as the highest in total consumption. Similarly, Figure 2 presents the visual depiction of countries around the world reaching their highest in 2010, followed by subsequent decrease.

------------------------------------------------------------------------

**Table 1**

Table \@ref(tab:Table1) displays the IQR, the range, Mean and Median values of total alcohol consumption in 2000, 2005, 2010, 2015 and 2018

```{r Table1, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}

# 2000 data
y2000 = newExtractedData[newExtractedData $ Year == 2000, ]

year2000Total = summary(y2000 $ Total_alcohol_consumption_per_capita)

year2000TotalArray = as.array(year2000Total)

year2000TotalData = data.frame(Statistics = rownames(year2000TotalArray), Values = year2000TotalArray, stringsAsFactors = FALSE)


# 2005 data
y2005 = newExtractedData[newExtractedData $ Year == 2005, ]

year2005Total = summary(y2005 $ Total_alcohol_consumption_per_capita)

year2005TotalArray = as.array(year2005Total)

year2005TotalData = data.frame(Statistics = rownames(year2005TotalArray), Values = year2005TotalArray, stringsAsFactors = FALSE)


# 2010 data
y2010 = newExtractedData[newExtractedData $ Year == 2010, ]

year2010Total = summary(y2010 $ Total_alcohol_consumption_per_capita)

year2010TotalArray = as.array(year2010Total)

year2010TotalData = data.frame(Statistics = rownames(year2010TotalArray), Values = year2010TotalArray, stringsAsFactors = FALSE)


# 2015 data
y2015 = newExtractedData[newExtractedData $ Year == 2015, ]

year2015Total = summary(y2015 $ Total_alcohol_consumption_per_capita)

year2015TotalArray = as.array(year2015Total)

year2015TotalData = data.frame(Statistics = rownames(year2015TotalArray), Values = year2015TotalArray, stringsAsFactors = FALSE)


# 2018 data
y2018 = newExtractedData[newExtractedData $ Year == 2018, ]

year2018Total = summary(y2018 $ Total_alcohol_consumption_per_capita)

year2018TotalArray = as.array(year2018Total)

year2018TotalData = data.frame(Statistics = rownames(year2018TotalArray), Values = year2018TotalArray, stringsAsFactors = FALSE)


# combining columns 
totalCombinedSummary = cbind(year2000TotalData$Statistics, year2000TotalData$Values.Freq, year2005TotalData$Values.Freq, year2010TotalData$Values.Freq, year2015TotalData$Values.Freq, year2018TotalData$Values.Freq)

# renaming columns 
colnames(totalCombinedSummary) = c("Statistics", "2000", "2005", "2010", "2015", "2018")

# presenting using kable()
kable(totalCombinedSummary, format = "html", align = "c", caption = "Alcohol Consumption Statistics throughout the Years") %>% kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```

------------------------------------------------------------------------

**Figure 1**

Figure \@ref(fig:Figure1) displays the trends in total alcohol consumption across geographical regions from 2000 to 2018, with the World's rate shown as a dot-dashed line

-   'Europe and Central Asia' region experienced a notable decrease in consumption, while 'Latin America and Caribbean' and 'Sub-Saharan Africa' regions saw slight declines

-   'East Asia and Pacific' and 'South Asia' regions presented rapid growth, particularly between 2005 and 2010

-   'North America' and 'Middle East and North Africa' regions had minor increases, however the 'Middle East and North Africa' remained the lowest

```{r Figure1, fig.cap = "Region Chart", echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}

# selecting specific regions 
specifiedRegions = c("World", "East Asia and Pacific", "Europe and Central Asia", "Latin America and Caribbean", "Middle East and North Africa", "North America", "South Asia", "Sub-Saharan Africa")

specifiedRegionsData = newExtractedData %>% filter(Entity %in% specifiedRegions)

# create chart with World being a dotdash line type 
regionChart = ggplot(specifiedRegionsData, aes(x = Year, y = Total_alcohol_consumption_per_capita, group = Entity, color = Entity)) + geom_line(aes(linetype= if_else(Entity == "World", "solid", "dotdash"))) + geom_point() + labs(x = "Year", y = "Alcohol Consumption in Litres", title = "Total Alcohol Consumption by Geographical Region") + scale_linetype_manual(values = c("solid", "dotdash"), guide = FALSE) + theme_minimal()

regionChart

```

------------------------------------------------------------------------

*Question 2*

-   How does the influence of religion relate to variations in total alcohol consumption per capita across different locations?

**Analysis**

Figure \@ref(fig:Figure1) highlighted the region 'Middle East and North Africa' as having had experienced the least growth and changes from 2000 to 2018 in comparison to the others. Notably, 'Middle East and North Africa' consistently exhibited the lowest alcohol consumption (consuming less than 1L each year). Similarly, Figure 2 shows the geographical locations: the Middle East, North Africa and the Greater Indonesia to have had minor changes over the years.

Table \@ref(tab:Table2) demonstrates consistency among the lowest 10 countries in alcohol consumption over the years, with only minor variations. Interestingly, a significant number of these countries are known for their devotion of Islamic values.

Although the impact of religion on alcohol consumption cannot be definitively determined without further global religious information. Figure \@ref(fig:Figure1), Table \@ref(tab:Table2) and Figure 2 suggest that these locations (the Middle East, North Africa and the Greater Indonesia) may experience lower alcohol consumption due to religious beliefs.

------------------------------------------------------------------------

**Figure 2**

Figure 2 presents the alcohol consumption per capita in 2000, 2005, 2010, 2015 and 2018. Some notable observations are:

-   In 2000, Russia showed high alcohol consumption but over time it declined to 11.19L in 2018

-   In 2010, most countries had their highest rate of alcohol consumption

-   From 2000 to 2018, the majority of North Africa, the Middle East and the Greater Indonesia regions presented minimal changes in alcohol consumption - this may be attributed by Islamic beliefs as the Quran deems drinking as haram

```{r, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}

# loading world map 
worldMap = map_data("world")

# selecting specific year 2000
extractedData2000 = filter(newExtractedData, Year == 2000)

worldMap $ iso3c = countrycode(worldMap$region, 'country.name', 'iso3c')

extractedData2000Map = left_join(worldMap, extractedData2000, by = c("iso3c" = "Code"))

# heat map based on total alcohol consumption in 2000 
dataMap2000 = ggplot() + geom_map(data = extractedData2000Map, map = extractedData2000Map, aes(x = long, y = lat, map_id = region, fill = Total_alcohol_consumption_per_capita), color = "black") + coord_equal() + theme_void() + scale_fill_gradient2(low = "white", mid = "lightblue", high = "darkred", name = "Alcohol Consumed (Litres)") + labs(title = "2000 Alcohol Consumption per Capita")


# selecting specific year 2005
extractedData2005 = filter(newExtractedData, Year == 2005)

extractedData2005Map = left_join(worldMap, extractedData2005, by = c("iso3c" = "Code"))

# heat map based on total alcohol consumption in 2005 
dataMap2005 = ggplot() + geom_map(data = extractedData2005Map, map = extractedData2005Map, aes(x = long, y = lat, map_id = region, fill = Total_alcohol_consumption_per_capita), color = "black") + coord_equal() + theme_void() + scale_fill_gradient2(low = "white", mid = "lightblue", high = "darkred", name = "Alcohol Consumed (Litres)") + labs(title = "2005 Alcohol Consumption per Capita")


# selecting specific year 2010
extractedData2010 = filter(newExtractedData, Year == 2010)

extractedData2010Map = left_join(worldMap, extractedData2010, by = c("iso3c" = "Code"))

# heat map based on total alcohol consumption in 2010 
dataMap2010 = ggplot() + geom_map(data = extractedData2010Map, map = extractedData2010Map, aes(x = long, y = lat, map_id = region, fill = Total_alcohol_consumption_per_capita), color = "black") + coord_equal() + theme_void() + scale_fill_gradient2(low = "white", mid = "lightblue", high = "darkred", name = "Alcohol Consumed (Litres)") + labs(title = "2010 Alcohol Consumption per Capita")


# selecting specific year 2015
extractedData2015 = filter(newExtractedData, Year == 2015)

extractedData2015Map = left_join(worldMap, extractedData2015, by = c("iso3c" = "Code"))

# heat map based on total alcohol consumption in 2015 
dataMap2015 = ggplot() + geom_map(data = extractedData2015Map, map = extractedData2015Map, aes(x = long, y = lat, map_id = region, fill = Total_alcohol_consumption_per_capita), color = "black") + coord_equal() + theme_void() + scale_fill_gradient2(low = "white", mid = "lightblue", high = "darkred", name = "Alcohol Consumed (Litres)") + labs(title = "2015 Alcohol Consumption per Capita")


# selecting specific year 2018
extractedData2018 = filter(newExtractedData, Year == 2018)   

extractedData2018Map = left_join(worldMap, extractedData2018, by = c("iso3c" = "Code"))

# heat map based on total alcohol consumption in 2018 
dataMap2018 = ggplot() + geom_map(data = extractedData2018Map, map = extractedData2018Map, aes(x = long, y = lat, map_id = region, fill = Total_alcohol_consumption_per_capita), color = "black") + coord_equal() + theme_void() + scale_fill_gradient2(low = "white", mid = "lightblue", high = "darkred", name = "Alcohol Consumed (Litres)") + labs(title = "2018 Alcohol Consumption per Capita")


# display graphs using par()
par(mfrow = c(3, 2))
plot(dataMap2000)
plot(dataMap2005)
plot(dataMap2010)
plot(dataMap2015)
plot(dataMap2018)

```

------------------------------------------------------------------------

**Table 2**

Table 2 \@ref(tab:Table2) presents the lowest 10 countries in 2000, 2005, 2010, 2015 and 2018 in alcohol consumption

```{r Table2, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}

# calculating the total alcohol consumption in 2000
total2000 = extractedData2000 %>% group_by(Entity) %>% summarize(Total = sum(Total_alcohol_consumption_per_capita))

# arranging total consumption in descending order
topTotal2000 = arrange(total2000, desc(Total))

# selecting the bottom 15 countries in 2000
tailTotal2000 = tail(topTotal2000, 10)


# calculating the total alcohol consumption in 2005
total2005 = extractedData2005 %>% group_by(Entity) %>% summarize(Total = sum(Total_alcohol_consumption_per_capita))

# arranging total consumption in descending order
topTotal2005 = arrange(total2005, desc(Total))

# selecting the bottom 15 countries in 2005
tailTotal2005 = tail(topTotal2005, 10)


# calculating the total alcohol consumption in 2010
total2010 = extractedData2010 %>% group_by(Entity) %>% summarize(Total = sum(Total_alcohol_consumption_per_capita))

# arranging total consumption in descending order
topTotal2010 = arrange(total2010, desc(Total))

# selecting the bottom 15 countries in 2010
tailTotal2010 = tail(topTotal2010, 10)


# calculating the total alcohol consumption in 2015
total2015 = extractedData2015 %>% group_by(Entity) %>% summarize(Total = sum(Total_alcohol_consumption_per_capita))

# arranging total consumption in descending order
topTotal2015 = arrange(total2015, desc(Total))

# selecting the bottom 15 countries in 2015
tailTotal2015 = tail(topTotal2015, 10)


# calculating the total alcohol consumption in 2018
total2018 = extractedData2018 %>% group_by(Entity) %>% summarize(Total = sum(Total_alcohol_consumption_per_capita))

# arranging total consumption in descending order
topTotal2018 = arrange(total2018, desc(Total))

# selecting the bottom 15 countries in 2018
tailTotal2018 = tail(topTotal2018, 10)


# combining the data for each year into one 
combinedData = bind_rows(mutate(tailTotal2000, Year = 2000), mutate(tailTotal2005, Year = 2005), mutate(tailTotal2010, Year = 2010), mutate(tailTotal2015, Year = 2015), mutate(tailTotal2018, Year = 2018))

# creating separate columns for each year
combinedData = pivot_wider(data = combinedData, id_cols = Entity, names_from = Year, values_from = Total)

# converting to character type and replacing NA values with '-'
combinedData = combinedData %>% mutate(across(everything(), as.character)) %>% replace_na(list(`2000` = '-', `2005` = '-', `2010` = '-', `2015` = '-', `2018` = '-'))

# display table using kable()
kable(combinedData, format = "html", align = "c", caption = "Bottom 10 Countries in Alcohol Consumption (Litres)") %>% kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```



# Introduction


## Gloria'section

### Research questions
Consuming too much alcohol and not being able to control drinking can harm an individual's physical or mental health and safety, causing to family or financial problems. Alcohol use disorders widely affect people across the globe. In this section, we will examine the death caused by alcohol use disorders __directly__ worldwide and attempt to answer the following two questions:

1. Which countries worldwide were most affected by alcohol use disorders?

2. Has the situation improved or got worse over the past ten years?


### Countries in cold areas tend to be more affected by alcohol use disorders. {.tabset}

Belarus and Mongolia had the highest __direct__ death rates (not including indirect death from suicide) of 21.80 and 17.05 per 100,000 people, respectively, followed by Russia, El Salvador, and Greenland (Table \@ref(tab:TableSummary1)).

Most of the countries with high death rates are in the subarctic or arctic zones in the northern hemisphere (Figure \@ref(fig:Visual1)). This is in line with findings from section 1. The cold and dark climates in these countries might have contributed to this observation.

However, countries in hot and humid weather such as Guatemala and Brazil also had high death rates from alcohol abuse. Countries in the southern subarctic zone with high alcohol consumption levels only showed low to moderate death rates. Therefore, cultural, genetic, historical, and religious factors cannot be ignored when investigating the underlying reason of high death rates in these countries.

#### Plot

```{r Visual1, fig.cap = "\\label{fig:Visual1}Annual Death rates from alcohol", echo = FALSE, warning = FALSE, message = FALSE}
# Load world map data
world<-map_data("world")

# Create data set and merge with map data
data_visual1 <- data_init_mod %>%
  rename("region" = "Entity") %>%
  filter(!is.na(Death_alcohol_use_disorders) & !Year %in% c("2000", "2005")) %>%
  group_by(region) %>%
  summarize(Mean = mean(Death_alcohol_use_disorders, na.rm = TRUE))

data_visual1 <- left_join(world, data_visual1, by = "region")

# Create heat map
visual_1 <- ggplot(data_visual1,aes(long,lat,group = group, fill = Mean, label = region)) +
  geom_polygon(color="white") +
  scale_fill_gradient(low="lightblue",high="red", name = "Death") +
  theme(panel.background = element_rect(fill = "lightgrey")) +
  xlab("") +
  ylab("Latitude") 

# Make plot interactive and add title and subtitle
visual_1_interactive <- ggplotly(visual_1) %>%
  layout(title = list(text = paste0("Death caused by alcohol use disorders",
                                    "<br>",
                                    "<sup>",
                                    "Average annual number of death from alcohol use disorder per 100,000 people",
                                    "</sup>")),
      annotations = list(x = 1, y = -0.1, text = "Data retreived from https://ourworldindata.org/alcohol-consumption.",
                         showarrow = F, xref='paper', yref='paper',
                         xanchor='right', yanchor='auto',
                         xshift=0, yshift=5,
                         font=list(size=10, color="black")))

visual_1_interactive

```

#### Table

```{r TableSummary1, echo = FALSE, warning = FALSE, message = FALSE}

data_table1 <- data_init_mod %>%
  filter(!is.na(Death_alcohol_use_disorders) & !Year %in% c("2000", "2005")) %>%
  group_by(Entity) %>%
  summarize(Mean = mean(Death_alcohol_use_disorders, na.rm = TRUE)) %>%
  arrange(desc(Mean)) %>%
  top_n(20)

# Make the table
table1 <- knitr::kable(data_table1,
                       digits = 2,
                       caption = "The top 20 countries with highest annual number of death caused by alcohol use disorders.")

table1
```


### The numbers fluctuate, but we have seen some improvements. {.tabset}

Table \@ref(tab:TableSummary2) shows the standard deviations of %death from alcohol per country from 2010-2019. Most countries with high variances also had high average %death rates such as Kazakhstan, Guatemala, Russia, Mongolia, etc.

In figure \@ref(fig:Visualtrend), rate groups are classified by the average annual %death rates over 2010-2019 into low (group1, <=1st quarter), medium (group2, <3rd quarter & >1st quarter) and high (group3, >=3st quarter). Only the top ten countries are shown in each group. __Note that the y axis are different in each subplot.__ This is to illustrate the trends in each group.

Zooming in onto the global trends, we can see that:

- Most countries with both medium and high average annual %death from alcohol showed some improvements with a decreased %death rate.

- Countries with low average annual %death from alcohol had more fluctuations and some even showed a slight increase.


#### Plot
```{r Visualtrend, fig.cap= "\\label{fig:Visualtrend}Changes of death rates from alcohol in selected countries 2010-2019", echo = FALSE, warning = FALSE, message = FALSE}
# Classfy groups depending on the mean %death over 2010-2019
data_rank <- data_init_mod %>%
  group_by(Entity) %>%
  summarize(Mean = mean(Death_alcohol_use_disorders, na.rm = TRUE)) %>%
  mutate(rategroup = case_when(Mean >=2.752 ~ "3",
                               Mean <2.752 & Mean >0.920 ~ "2",
                               Mean <=0.920 ~ "1"))

# Create list of countries with high, medium, low %death rates
data_rank_low <- data_rank %>%
  filter(rategroup == "1") %>%
  arrange(desc(Mean))

data_rank_medium <- data_rank %>%
  filter(rategroup == "2") %>%
  arrange(desc(Mean))

data_rank_high <- data_rank %>%
  filter(rategroup == "3") %>%
  arrange(desc(Mean))

# Classify entities by 
data_rank2 <- data_init_mod %>%
  group_by(Entity) %>%
  mutate(rategroup = case_when(Entity %in% data_rank_high$Entity ~ "3",
                               Entity %in% data_rank_medium$Entity ~ "2",
                               Entity %in% data_rank_low$Entity ~ "1"))

# Create a line graph to show global trends
visual2 <- data_rank2 %>%
  filter(Year > "2009") %>%
  filter(Entity %in% head(data_rank_high$Entity, 10)|Entity %in% head(data_rank_medium$Entity, 10)|Entity %in% head(data_rank_low$Entity, 10)) %>%
  filter(!is.na(Death_alcohol_use_disorders)) %>%
  ggplot(aes(x = as.character(Year), y = Death_alcohol_use_disorders, group = Entity, color = rategroup)) +
  geom_line() +
  scale_color_manual(values=c("#FAA19B", "#f66257", "#b73229"))+
  facet_grid(rategroup~., scales = "free") +
  xlab("") +
  ylab("%Death from alcohol")

visual_2_interactive <- ggplotly(visual2) %>%
  layout(title = list(text = paste0("Death changes from alcohol use disorders 2010-2019 per group",
                                    "<br>",
                                    "<sup>")),
         annotations = list(x = 1, y = -0.1, text = "Data retreived from https://ourworldindata.org/alcohol-consumption.",
                         showarrow = F, xref='paper', yref='paper',
                         xanchor='right', yanchor='auto',
                         xshift=0, yshift=0,
                         font=list(size=10, color="Black")))

visual_2_interactive

```

#### Table
```{r TableSummary2, echo = FALSE, warning = FALSE, message = FALSE}
data_table2 <- data_init_mod %>%
  filter(!is.na(Death_alcohol_use_disorders) & !Year %in% c("2000", "2005")) %>%
  group_by(Entity) %>%
  summarize(Standard_deviation = sd(Death_alcohol_use_disorders, na.rm = TRUE)) %>%
  arrange(desc(Standard_deviation)) %>%
  top_n(20)

# Make the table
table2 <- knitr::kable(data_table2,
                       digits = 2,
                       caption = "The top 20 countries with highest variance in number of death caused by alcohol use disorders.")

table2

```




## Tony's section
```{r}
data <- read.csv("Merge_all_2010_2020_updated.csv")

alc_sex_regions <- data %>%
  select(Entity, Year, Prevalence_alcohol_use_disorders_male, 
         Prevalence_alcohol_use_disorders_female) %>%
  na.omit() %>%
  filter(Entity %in% c("African Region (WHO)", "Australia", "China", 
                       "European Region (WHO)", "United Kingdom", 
                       "United States"))

alc_sex_income <- data %>%
  select(Entity, Year, Prevalence_alcohol_use_disorders_male, 
         Prevalence_alcohol_use_disorders_female) %>%
  na.omit() %>%
  filter(Entity %in% c("World Bank Low Income", "World Bank Lower Middle Income", 
                       "World Bank Upper Middle Income", "World Bank High Income"))
```





# What are the variations in the prevalence of alcohol disorders among men and women across different income groups?

```{r, echo=FALSE}
male_income <- ggplot(data = alc_sex_income, aes(x = Year)) +
  geom_line(aes(y = Prevalence_alcohol_use_disorders_male, color = Entity)) +
  labs(x = "Year", y = "% with alcohol use disorders", title = "Male", color = "Income Group") +
  scale_color_manual(values = c("blue", "green", "orange", "red"),
                     labels = c("Low", "Lower Middle", "Upper Middle", "High"),
                     breaks = c("World Bank Low Income", "World Bank Lower Middle Income", 
                                "World Bank Upper Middle Income", "World Bank High Income")) +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = seq(2010, 2019, 2)) + 
  scale_y_continuous(breaks = seq(1.75, 3, 0.25))

female_income <- ggplot(data = alc_sex_income, aes(x = Year)) +
  geom_line(aes(y = Prevalence_alcohol_use_disorders_female, color = Entity)) +
  labs(x = "Year", y = "% with alcohol use disorders", title = "Female", color = "Income Group") +
  scale_color_manual(values = c("blue", "green", "orange", "red"),
                     labels = c("Low", "Lower Middle", "Upper Middle", "High"),
                     breaks = c("World Bank Low Income", "World Bank Lower Middle Income", 
                                "World Bank Upper Middle Income", "World Bank High Income")) +
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = seq(2010, 2019, 2)) +
  scale_y_continuous(breaks = seq(0.4, 1.6, 0.2))

grid.arrange(male_income, female_income, ncol = 1)
```

```{r, echo=FALSE}
alc_sex_income_10_19 <- alc_sex_income %>%
  filter(Year %in% c(2010,2019))

kable(alc_sex_income_10_19, caption = "% of Males and Females with Alcohol Use Disorders in 2010 and 2019")
```

An association between income and alcohol disorders reveals noteworthy deviations from 2010 to 2019 but it is also important to consider the influence of other socioeconomic factors.
There is a noticeable distinction between the high-income group and other income groups. The percentage of alcohol disorders is significantly greater or approximately doubled when compared with other income groups. Females in the high-income group may be subject to societal expectations or gender roles which lead to increased alcohol related issues. 
There is a consistent trend indicating a higher prevalence of alcohol disorders among individuals from the higher income groups possibly with work pressures. Interestingly, the income groups among both males and females both have the lower-middle income group with the lowest rate of alcohol disorders. This may be due to considerations such as lower financial stressors and mental and physical well-being. However, the low and upper-middle income groups have a higher rate particularly among the males. Similarly, this may be due to financial and work stressors. 

# What is the temporal trend in the prevalence of alcohol disorders among men and women across major regions?

```{r, echo=FALSE}
ggplot(data = alc_sex_regions, aes(x = Year)) +
  geom_line(aes(y = Prevalence_alcohol_use_disorders_male, color = "Male")) +
  geom_line(aes(y = Prevalence_alcohol_use_disorders_female, color = "Female")) +
  labs(x = "Year", y = "% with alcohol use disorders", 
       title = "% of Alcohol Disorders of each Major Region", 
       color = "Sex") +
   scale_color_manual(values = c("blue", "hot pink"),
                     labels = c("Male", "Female"),
                     breaks = c("Male", "Female")) + 
  theme(legend.position = "right", plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(breaks = seq(2010, 2019, 2)) + 
  facet_wrap(~ Entity, ncol = 2, scales = "free_y")

```

```{r, echo=FALSE}
alc_sex_regions_10_19 <- alc_sex_regions %>%
  filter(Year %in% c(2010,2019))

kable(alc_sex_regions_10_19, caption = "% of Males and Females with Alcohol Use Disorders in 2010 and 2019")
```

Across all regions, it is clear there is a higher percentage of males with alcohol disorders than females and this has not changed significantly over the period 2010 to 2019. However, there is no evident similar trends between each region as there are various attributable factors such as evolving social norms, increased alcohol availability and alterations in cultural traditions. Regardless, it is apparent over this period, males with alcohol issues are increasing in Australia, China and the United Kingdom. 

